## Clustering  

Libraries  
```{r}
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(cluster) #algorithms for clustering
library(factoextra) #visualization

```

Read in data  
```{r}
customers = read_csv("CustomerData.csv")
str(customers)
summary(customers)
```

Preparing the data. Remove missingness (there is none in this data) or impute missing values.  

We also scale the data. This is critical for quantitative data to ensure that no variable (particularly a variable with large values, skews the data and the resulting clusters).  
```{r}
customers_scaled = scale(customers) 
summary(customers_scaled)
#scale works by calculating the mean and standard deviation of the entire variable, then scales each element by subtracting the mean and dividing by the standard deviation  
```

Perform k-means clustering with a pre-specified number of clusters. I selected 3 clusters. Note that k-means uses randomness, so we need to use set.seed to ensure same clusters.   
```{r}
set.seed(1234)
clusters1 <- kmeans(customers_scaled, 3)
clusters1 #don't do this for very large datasets as the cluster for each row is shown
```

Visualize the clustering  
```{R}
fviz_cluster(clusters1, customers_scaled)
```

Visually identify optimal number of clusters  
```{r}
set.seed(123)
fviz_nbclust(customers_scaled, kmeans, method = "wss") #minimize within-cluster variation
```
Another method  
```{r}
set.seed(123)
fviz_nbclust(customers_scaled, kmeans, method = "silhouette") #maximize how well points sit in their clusters
```

Let's try 5 clusters  
```{r}
set.seed(1234)
clusters2 <- kmeans(customers_scaled, 5)
clusters2 
fviz_cluster(clusters2, customers_scaled)
```





